---
title: "CodaBonito"
author: "Aditya Kothari"
date: "`r Sys.Date()`"
output: prettydoc::html_pretty
vignette: >
  %\VignetteIndexEntry{Introduction to CodaBonito}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r Setup, cache=FALSE, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 15, fig.height = 8, results = 'hide'}

library(knitr)
read_chunk('./CodaBonito.R')

```

```{r SetupChunk, echo = F}
```

This document offers a brief introduction to the functions in this library. 
Some function might be fairly straightforward to create but some functions may 
need you to look up the help entry before trying it out.

## Disclaimer / Call for Inputs

This library is very much in development. It is a compilation of code I've 
written across different project which is why there may be syntactical 
inconsistencies, such as some functions using player names and some others 
using player IDs as a reference for players. Some of this is also code that 
I have never published the output of so it may need additional arguments. I
have, however, tried to give enough documentation for each function so anyone
trying to use the library should be well equipped with instructions and a basic
understanding of how the functions work.

Any inputs and feedback is welcome. If you're on github, then head to
github.com/thecomeonman/CodaBonit, or else get in touch <a href = 
"twitter.com/thecomeonman">on Twitter</a>

## How to get started

Install R from https://cran.r-project.org

Open R and run this command in the console - 
`install.packages("devtools")`
`library(devtools)`
`install_github("thecomeonman/CodaBonito")`

And you're ready to run the examples below!

## Data

I have added some fake data along with the package to be able to better 
explain the usage of these functions.

`dtPlayerMetrics` - aggregated data for players is typically in a format 
similar to this, with some extra details about the team they play for, their 
age, etc.
```{r DataDescriptionPlayerMetrics, echo = F}
```

`dtMetricCategorisation` - some metadata about the metrics. 
- variableLabel is the name that will be displayed in charts for that metric, 
- variableCategory is the grouping of variables used in some visualisations,
like `fNormalisedValueChart`
- HighValueIsBad is marked true for variables where a high value is bad. 
Variables such as fouls and goals conceded would be true.
```{r DataDescriptionMetricCategorisation, echo = F}
```

`dtPasses` - passing data. 
- x,y denote the start coordinates of the pass
- endX, endY denote the end coordinates of the pass
- passLength is the length of the pass
- passAngle is the angle of the pass in radians  ( 180 degress = pi radians )
where 0 is along the pitch from defense to offense.
- Success 1 for successful pass, 0 for failed pass

```{r DataDescriptionPasses, echo = F}
```

`dtFormation` - formation coordinates
```{r DataDescriptionFormation, echo = F}
```

`dtPlayerLabels` - player labels
```{r DataDescriptionPlayerLabels, echo = F}
```

## Visualisations

### fAddPitchLines

fAddPitchLines draws pitch markings with further customisations available

```{r fAddPitchLines}
```

You can add whatever stats you want on top of it like regular ggplot2

```{r fAddPitchLinesData}
```

### theme_pitch

If you aren't interested in having the axis markings, etc., use theme_pitch

```{r theme_pitch}
```

### fNormalisedValueChart
```{r fNormalisedValueChart}
```

### fPercentileBarChart
```{r fPercentileBarChart}
```

### fPercentileBarChart with AbsoluteIndicator

Percentiles can be a little misleading if the underlying numbers aren't uniformly distributed. You can add annotations for an indicator of the absolute spread of the values and where this particular player's values fall within that spread.

```{r fPercentileBarChartAbsoluteIndicator}
```

### fRadarPercentileChart

I disapprove of radar charts. It's a bad visualisation, prone to 
misinterpretation. They seem to be the accepted norm of comparing players 
though which is why I had to sell out and have an implementation of that in the 
package, but I've added a warning which states how you should use 
fPercentileBarChart instead as that is a better structured visualisation.
```{r fRadarPercentileChart}
```

### fPlotSonar
```{r fPlotSonar}
```

### fPassNetworkChart
```{r fPassNetworkChart}
```

### fXgBuildUpComparison
```{r fXgBuildUpComparison}
```


### fDrawVoronoi

WIP using the data structure from https://github.com/metrica-sports/sample-data

```{r fDrawVoronoi}



```

## Logic and Algorithms

### fEMDDetailed
A function to calculate earth mover's distance. It offers more flexibility and
transparency than emdist:emd.

Any distance matrix can be used to calculated EMD, but emdist:emd insists on 
getting the raw distributions with only up to four dimensions. fEMDDetailed
only requires a distance matrix between each combination of observations in the 
two datasets, irrespective of the nature of the data.

```{r fEMDDetailed}
```

On the topic of transparency, one of the things I find very useful is that you 
can now see how much distance is being contributed by each observation.

```{r fEMDDetailedExtra}
```

## Data Parsing

You will find `fJsonToListOfTables`, `fJsonToTabular`, `fParseTrackingData` useful. They aren't glamarous enough to be demoed here but the documentation should help you use those functions.

